Kernel Phase and Coronagraphy with Automatic Differentiation

### 2020-05-18 ms.tex
Updates from Overleaf

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
% optical imaging with aberrations
Aberrations in the optical path introduce distortions and speckles in the resulting images which limit the performance of cameras at high angular resolution.
% self cal - closure phase and kernel phase
Self-calibrating observables such as the `closure phase' or `bispectrum' have been widely used in optical and radio astronomy to correct for this: a linear combination of measurements with phase noise can be made such that error terms cancel, leaving a subspace of high signal-to-noise observables. Kernel phases are a generalization of closure phases, in which the effect of aberration on an arbitrary pupil is approximated as a linear phase transfer matrix. 
%coronagraphs are hard
Finding this operator is analytically tractable between a single pupil and focus but is much more difficult for other optical systems of interest, such as coronagraphs.
% autodiff can make this practical for arbitrary optical systems
Automatic differentiation or `backpropagation' software now makes calculating derivatives with respect to aberrations in arbitrary planes straightforward for any optical system.
% we demonstrate a simple example of the kernel of a monochromatic coronagraph
We reproduce existing kernel phase theory within this framework and demonstrate an extension to the case of a Lyot coronagraph, which is found to have self-calibrating combinations of speckles which are resistant to phase noise.
%phase design
We present a Python package \textsc{morphine}, with an interface similar to the popular package \textsc{poppy}, for optical simulation with automatic differentiation. These methods may be useful for designing improved astronomical optical systems by gradient descent.
\href{https://github.com/benjaminpope/morphine}{\color{linkcolor}\faGithub} % shamelessly borrowed from Luger!

@@

Close to the topic of this paper, autodiff has been applied to the problem of phase retrieval \citep{jurling14,paine19}, inferring a wavefront from a PSF. In this context, a key advance of autodiff over previous methods is that we can trivially account for pixel sampling/binning and detector nonlinearity. These will be important issues when considering that the JWST mid-infrared imager MIRI will have significant detector nonlinearity \citep{rieke15}, or where we may wish to look at saturated sources.  While they do not address more complex optical systems, this method may be straightforwardly applicable to sensing non-common-path errors in coronagraphic images. While \citet{jurling14} derive analytic expressions for many optically-relevant operations, machine-learning software that has become available since then have significantly widened the range of options and introduced more user-friendly APIs. These approaches have been taken up outside of optical imaging, for example in X-ray coherent diffractive imaging \citep{kandel19,nashed19}.

@@

We then use Markov Chain Monte Carlo \citep{metropolis53} to sample from the posterior assuming Gaussian likelihoods, using the \texttt{emcee} affine-invariant ensemble sampler \citep{emcee}. We obtain binary parameters of separation $\rho = 115.3 \pm 1.2$~mas, position angle $\theta = 85.6 \pm 0.27$~degrees, and contrast $c = 26.3 \pm 0.6$: these are close to the \citet{martinache20} figures, but differ by much more than their small uncertainties. For comparison, the contrast ratio of the system is expected from resolved imaging to be $28.0 \pm 8.3$ in $K$-band \citep{hinkley11}.

\begin{figure}
\plotone{comparison_posterior.pdf}
\caption{Posterior distributions for the binary parameters of $\alpha$~Ophiuchi as determined by the autodiff pipeline (black) and the \citet{martinache20} implementation (blue). Corner plots produced by \texttt{corner.py} \citep{corner}. \label{colinearity}}
\end{figure}

In order to investigate this discrepancy, we generated a grid of forwards models

>>>>

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
% optical imaging with aberrations
Aberrations in the optical path introduce distortions and speckles in the resulting images which limit the performance of cameras at high angular resolution.

% self cal - closure phase and kernel phase
Self-calibrating observables such as the `closure phase' or `bispectrum' have been widely used in optical and radio astronomy to correct for this: a linear combination of measurements with phase noise can be made such that error terms cancel, leaving a subspace of high signal-to-noise observables. Kernel phases are a generalization of closure phases, in which the effect of aberration on an arbitrary pupil is approximated as a linear phase transfer matrix.

%coronagraphs are hard
Finding this operator is analytically tractable between a single pupil and focus but is much more difficult for other optical systems of interest, such as coronagraphs.
% autodiff can make this practical for arbitrary optical systems
Automatic differentiation or `backpropagation' software now makes calculating derivatives with respect to aberrations in arbitrary planes straightforward for any optical system.

% we demonstrate a simple example of the kernel of a monochromatic coronagraph
We reproduce existing kernel phase theory within this framework and demonstrate an extension to the case of a Lyot coronagraph, which is found to have self-calibrating combinations of speckles which are resistant to phase noise, but only in the very high wavefront quality regime. We also use this framework to reanalyze Palomar adaptive optics observations of the binary $\alpha$~Ophiuchi, finding broad consistency and slight systematics between the new pipeline and the existing standard.

%phase design
We present a Python package \textsc{morphine}, with an interface similar to the popular package \textsc{poppy}, for optical simulation with automatic differentiation. These methods may be useful for designing improved astronomical optical systems by gradient descent.
\href{https://github.com/benjaminpope/morphine}{\color{linkcolor}\faGithub} % shamelessly borrowed from Luger!

@@

Close to the topic of this paper, autodiff has been applied to the problem of phase retrieval \citep{jurling14,paine19}, inferring a wavefront from a PSF. In this context, a key advance of autodiff over previous methods is that we can trivially account for pixel sampling/binning and detector nonlinearity. These will be important issues when considering that the JWST mid-infrared imager MIRI will have significant detector nonlinearity \citep{rieke15}, or where we may wish to look at saturated sources.  While they do not address more complex optical systems, this method may be straightforwardly applicable to sensing non-common-path errors in coronagraphic images. While \citet{jurling14} derive analytic expressions for many optically-relevant operations, machine-learning software that has become available since then have significantly widened the range of options and introduced more user-friendly APIs. These approaches have been taken up outside of optical imaging, for example in X-ray coherent diffractive imaging \citep{kandel19,nashed19} and nanotomography \citep{Dueaay3700}.

@@

We then use Markov Chain Monte Carlo \citep{metropolis53} to sample from the posterior assuming Gaussian likelihoods, using the \texttt{emcee} affine-invariant ensemble sampler \citep{emcee}. We obtain binary parameters of separation $\rho = 115.3 \pm 1.2$~mas, position angle $\theta = 85.6 \pm 0.27$~degrees, and contrast $c = 26.3 \pm 0.6$: these are close to the \citet{martinache20} figures, but differ by much more than their small uncertainties. For comparison, the contrast ratio of the system is expected from resolved imaging to be $28.0 \pm 8.3$ in $K$-band \citep{hinkley11}. Posteriors for both inferences are shown in Figure~\ref{comparison_posterior}.

\begin{figure}
\plotone{comparison_posterior.pdf}
\caption{Posterior distributions for the binary parameters of $\alpha$~Ophiuchi as determined by the autodiff pipeline (black) and the \citet{martinache20} implementation (blue). Corner plots produced by \texttt{corner.py} \citep{corner}. \label{comparison_posterior}}
\end{figure}

In order to investigate this discrepancy, we generated a grid of forwards models. Using the same code used to produce the kernel phase autodiff model, we generated input wavefronts with 20 Zernike modes of aberrations, each with random amplitudes normally-distributed with 15\,nm standard deviation, for a total peak-to-peak amplitude of 206\,nm. These were then shifted and added to simulate a range of binaries at different configurations and contrasts, and both the \citet{martinache20} pipeline and the autodiff pipeline used to retrieve their parameters. The results are shown in Figure~\ref{method_comparison_ensemble}. The pipelines achieve broadly comparable performance on these simulations, with the \citet{martinache20} method being slightly better at closer separations and higher contrasts and autodiff performing marginally better at larger separations. For simulated data with fiducial $\rho = 125$\,mas, $\theta=86^\circ$, and $c = 25$, there is no sign of the systematic difference seen in real data. In \citet{martinache20} a large error term was added in quadrature in order to account for additional noise in the data unexplained by the diversity over exposures and uncalibrated by the kernel phase model. In order to achieve $\chi^2 = 1$ for simulated data, an order of magnitude less additional error was needed in quadrature using both models.

\begin{figure}
\plotone{method_comparison_ensemble.pdf}
\caption{Comparison of \citet[][blue stars]{martinache20} and autodiff (orange dots) kernel phase pipelines on simulated data, as a function of input contrast (left) and separation (right). When inspected, the samples with worse separation error after both pipelines are those at higher contrasts. Uncertainties are omitted for clarity, and because they are poorly defined in this context. \label{method_comparison_ensemble}}
\end{figure}

In numerical experiments, subpixel mismatch in Fourier transform coordinates or even Fourier centering convention could cause mismatches between simulation paramet and retrievals at around the scale of the mismatch between the two methods.  These results are in line with the \citet{martinache_habilitation} conclusion that inexact Fourier sampling contributes significant systematic noise to kernel phases.The final version used for the autodiff pipeline uses the exact optical model used to generate the data, and a differentiable version of the exact $u,v$ transform implementation used in \textsc{xara}; we do not believe this to be the main cause of the small discrepancy between autodiff and conventional parameters, but it does illustrate that model mis-specification may be a contributing factor.
